Cluster Difference Imaging Photometric Survey. III. An Open Cluster With a 250\,pc Halo

### 2021-02-03 cluster_paper/ms.tex
update paper

\newcommand{\nkinematic}{3{,}018\ } % "fullfaint" kinematic sample, core+halo. from any call of earhat.helpers._get_fullfaint_dataframes
\newcommand{\nautorot}{1{,}276\ } % drivers.plot_full_kinematics_X_rotation; 2021.01.29

@@

% Context, aims, methods, results, conclusions.
\begin{abstract}
  Recent analyses of the Gaia data have reported the existence of
  diffuse stellar populations surrounding nearby open clusters.  The
  stars in these ``halos'' could have escaped from the cores of the
  cluster, or they could be remnants of a star formation complex that
  has since dispersed.  They could also be field stars---false
  positives from overly aggressive clustering algorithms.  In this
  study, we focus on a 250\,pc halo reported
  around the $\approx$120\,Myr open cluster NGC\,2516.  The classical
  tidal radius of the cluster is {\bf $\sim$XX\,pc}.  Using photometry
  from Gaia, rotation periods from TESS, and lithium measurements from
  Gaia-ESO and GALAH, we find that the halo of NGC\,2516 is real and
  coeval with the core.  Two in three kinematically-selected halo
  members 25-250\,pc from the cluster center show rotation periods
  consistent with a gyrochronological age of 120\,Myr.  Stars in the
  inner 25\,pc show higher period detection rates than the
  halo (73\% vs.\ 66\%), but both vastly exceed our field star
  comparison sample ({\bf XX\%}).  The rotation rate and lithium abundance
  are also anti-correlated at fixed stellar mass, as is the case with
  comparable clusters.  This work expands the set of confirmed
  NGC\,2516 members by a factor of $\approx$2, and quantifies the
  degree to which Gaia-based analyses can identify dispersed stellar
  populations against a background of field star contaminants.  We
  highlight the implications for spectroscopic survey targeting, open
  cluster dispersal, and planet searches around young stars.
\end{abstract}
% Rotators < 25pc : 0.72929 +/- 0.00013
% Rotators > 25pc : 0.66172 +/- 0.00023

@@

In this work, we also define a set of nearby field stars in the
``neighborhood'' of \cn.  Based on the observed distribution of halo
members, we draw these stars randomly from the following ``cube'' in
right ascension, declination, and parallax:
\begin{align}
  \alpha\,[\mathrm{deg}] &\in [108, 132], \\
  \delta\,[\mathrm{deg}] &\in [-76, -45], \\
  \pi\,[\mathrm{mas}] &\in [1.5, 4.0].
\end{align}
We imposed a magnitude limit of $G=19$\,mag, and ran the queries
using the \texttt{astroquery.gaia} module (CITE).
% TODO: why?
We allowed the number of stars in the comparison sample to exceed that
in the cluster sample by a factor of a few, to ensure broad sampling
of stellar masses and evolutionary states.
We also required the comparison sample to not overlap with the cluster
sample, which led to the omission of 1.1\% of the stars drawn over the
volume noted above.

@@

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\subfloat{
			\includegraphics[width=0.75\textwidth]{f2a.pdf}
		}

		\subfloat{
			\includegraphics[width=0.75\textwidth]{f2b.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf HR diagrams of NGC\,2516, using Gaia EDR3 photometry.}
    {\it Top:} The core (black) shows a clean sequence consistent with
    stars with a fixed age and metallicity, and varying mass.  The
    halo (blue) is similar, but somewhat noisier.  The faintest M
    dwarfs in the core and halo are brighter than in the field star
    comparison sample (gray), consistent with these stars having not
    yet reached the ZAMS.
    {\it Bottom:} Reported members of the halo, as a function of
    galactic latitude. Can the additional scatter in the halo be
    understood through differential reddening?
    {\bf Maybe}.
    \label{fig:hr}
  }
\end{figure*}

@@

Figure~\ref{fig:hr} presents similar results to what these
investigators have already found.  The core members of the cluster
show a clean sequence consistent with stars with a fixed age and
metallicity, and varying mass.
The halo members are roughly consistent with this, but they do show
greater scatter.  One possible explanation for this scatter is that
the halo is more contaminated by field stars.
Another, explored in the other panel of the figure, is differential
reddening.
The halo is reported to span 20$^\circ$ on-sky, and varies in position from
about $b=-12^\circ$ to $b=-20^\circ$, with the stars closest to the
galactic plane also being further from the Sun
by up to 200\,pc (Figure~\ref{fig:gaia6d}).

@@

@@

\section{Rotation $\otimes$ RUWE}

>>>>

\newcommand{\nautorot}{1{,}276\ } % drivers.plot_full_kinematics_X_rotation; 2021.01.29
\newcommand{\nkinematic}{3{,}018\ } % "fullfaint" kinematic sample, core+halo. from any call of earhat.helpers._get_fullfaint_dataframes

\newcommand{\ncalibration}{9{,}619{,}784} % /nfs/phtess2/ar0/TESS/PROJ/lbouma/CDIPS_LCS/counts/check_sum, 2021.02.02
\newcommand{\nnbhd}{13{,}843\ } % make_compstar_NGC_2516_sourcelist.py
\newcommand{\nnbhdcaliblc}{10{,}314\ } % make_compstar_NGC_2516_sourcelist.py
\newcommand{\nnbhdcalibstar}{2{,}212\ } % make_compstar_NGC_2516_sourcelist.py

@@

% Context, aims, methods, results, conclusions.
\begin{abstract}
  Recent analyses of the Gaia data have reported the existence of
  diffuse stellar populations surrounding nearby open clusters.  The
  stars in these ``halos'' could have escaped from the cores of the
  cluster, or they could be remnants of a star formation complex that
  has since dispersed.  They could also be field stars---false
  positives from overly aggressive clustering algorithms.  In this
  study, we focus on a 250\,pc halo reported
  around the $\approx$120\,Myr open cluster NGC\,2516.  The classical
  tidal radius of the cluster is {\bf $\sim$XX\,pc}.  Using photometry
  from Gaia, rotation periods from TESS, and lithium measurements from
  Gaia-ESO and GALAH, we find that the halo of NGC\,2516 is real and
  coeval with the core.  Two in three kinematically selected halo
  members 25-250\,pc from the cluster center show rotation periods
  consistent with a gyrochronological age of 120\,Myr.  Stars in the
  inner 25\,pc show higher period detection rates than the
  halo (73\% vs.\ 66\%), but both vastly exceed our field star
  comparison sample ({\bf XX\%}).  The rotation rate and lithium abundance
  are also anti-correlated at fixed stellar mass, as is the case with
  comparable clusters.  This work expands the set of confirmed
  NGC\,2516 members by a factor of $\approx$2, and quantifies the
  degree to which Gaia-based analyses can identify dispersed stellar
  populations against a background of field star contaminants.  We
  highlight the implications for spectroscopic survey targeting, open
  cluster dispersal, and planet searches around young stars.
\end{abstract}
% Rotators < 25pc : 0.72929 +/- 0.00013
% Rotators > 25pc : 0.66172 +/- 0.00023

@@

In this work, we also define a set of nearby field stars in the
``neighborhood'' of \cn.  Based on the observed distribution of halo
members, we draw these stars randomly from the following ``cube'' in
right ascension, declination, and parallax:
\begin{align}
  \alpha\,[\mathrm{deg}] &\in [108, 132], \\
  \delta\,[\mathrm{deg}] &\in [-76, -45], \\
  \pi\,[\mathrm{mas}] &\in [1.5, 4.0].
\end{align}
We imposed a magnitude limit of $G=19$\,mag, and ran the queries
using the \texttt{astroquery.gaia} module (CITE).
% TODO: why?
We allowed the number of stars in the comparison sample to exceed that
in the cluster sample by a factor of $\approx$5, to ensure broad sampling
of stellar masses and evolutionary states.
We also required the comparison sample to not overlap with the cluster
sample, which led to the omission of 1.1\% of the stars drawn over the
volume noted above.

@@

\begin{figure*}[t]
	\begin{center}
		\leavevmode
			\includegraphics[width=0.75\textwidth]{f2a.pdf}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf HR diagram of NGC\,2516, using Gaia EDR3 photometry.}
    The core (black) shows a clean sequence consistent with
    stars with a fixed age and metallicity, and varying mass.  The
    halo (blue) is similar, though somewhat noisier.  The faintest M
    dwarfs in the core and halo are brighter than in the field star
    comparison sample (gray), consistent with these stars having not
    yet reached the ZAMS.
    \label{fig:hr}
  }
\end{figure*}

@@

Figure~\ref{fig:hr} presents similar results to what these
investigators have already found.  The core members of the cluster
show a clean sequence consistent with stars with a fixed age and
metallicity, and varying mass.  The halo members show a roughly
consistent main sequence, but with somewhat greater scatter.  One
possible explanation for this scatter is that the halo is more
contaminated by field stars.  Another possible explanation is
differential reddening.  The halo is reported to span 20$^\circ$
on-sky, and varies in position from about $b=-12^\circ$ to
$b=-20^\circ$, with the stars closest to the galactic plane also being
further from the Sun by up to 200\,pc (Figure~\ref{fig:gaia6d}).
Empirical evidence for differential reddening is discussed further in
Appendix~\ref{app:diffred}. A priori, both field star contamination
and differential reddening could well play a role.

@@

\paragraph{Comparison sample}
To provide a basis for comparison, we also opted to search the
``calibration'' light curves ($Rp<13$) that were created as a part of
the CDIPS project. 
Over the southern sky (Sectors 1-13 of TESS), this corresponded to a sample of \ncalibration stars.
Cross-matching these against \nnbhd randomly drawn stars in the
neighborhood of \cn yielded \nnbhdcalibstar unique stars, with
\nnbhdcaliblc sector-specific TESS light curves.

@@

\section{Differential Reddening}
\label{app:diffred}

\begin{figure*}[t]
	\begin{center}
		\leavevmode
			\includegraphics[width=0.75\textwidth]{f2b.pdf}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf HR diagram of NGC\,2516, using Gaia EDR3 photometry.}
    {\it Bottom:} Reported members of the halo, as a function of
    galactic latitude. Can the additional scatter in the halo be
    understood through differential reddening?
    {\bf Maybe}.
    \label{fig:diffred}
  }
\end{figure*}

Figure~\ref{fig:diffred}.

\section{Rotation $\otimes$ RUWE}
\label{app:ruwe}
